Succinct representations for abstract interpretation

### 2011-10-29 papers/PLDI2012/implicitization.tex

We experimented with a heuristic
of dynamic construction of the $\sigma$ function, adapted from
\citet{DBLP:conf/pldi/GulwaniZ10}. 
\DM{Cette heuristique, je croyais que c'était nous qui l'avions trouvée,
et qu'elle n'était pas dans Gulwani?}
We start with one single
disjunct per control point ($m=1$) and an empty map $\sigma$.
$M$ denotes an upper bound on the number of disjuncts per control point.
$\sigma(j,k)$ is then constructed on the fly only when needed.
When $\sigma(j,k)$ needs to be computed, we first compute the image of the
abstract value $X_k$ by the path indexed by $j$, and try to find an existing
disjunct of index $k'$ so that the least upper bound of the two abstract values is exactly their union (this can be tested using SMT-solving).
If such an index exists, then we set $\sigma(j,k) = k'$.
Otherwise:
\begin{itemize}[itemsep=-1\parsep,topsep=0pt]
	\item if $m < M$, we increase $m$ by $1$ and define $\sigma(j,k) = m$
	\item if $m = M$, we define $\sigma(j,k) = M$
\end{itemize}

@@

\DM{Julien: pour la guided analysis, a priori ça arrive souvent cette histoire d'edge pas prise car non encore prenable?}
In the event that a node is reachable only by a single control-flow edge (which may occur because of dead code, or during the first phases of guided static analysis), the $\phi$ operation reduces to a copy of the values flowing from that edge. In this case, our tool just propagates symbolic values through the predecessor node, without introducing $\phi$-variables.

@@

Statistically, the domain of convex polyhedra gives the better results, but
commonly yields greater invariants than the domains of octagons/intervals;
this is a known weakness of its widening operator.
\DM{Ici on pourrait citer Monniaux+Le Guen si c'etait publie!}
The Octagon domain appears to be much better than intervals; this is
unsurprising since in most programs and libraries, bounds on loop indices are
non constant: they depend on some parameters (array sizes etc.).
\JH{Commenter les resultats}

@@

The analysis of linear equalities (PKEQ) performs very fast compared
to other abstract domains, but yields very imprecise invariants:
it only detects relations of the form $\sum_i a_i x_i = C$ where $a_i$ and $C$ are constants.
The comparison between the domain of convex polyhedra and linear congruences
shows the percentage of points where the use of congruences refines the
invariant. This is the case for $1.04\%$ of the control points in $P_R$.
\DM{Je croyais qu'on ne pouvait pas utiliser PPLGRID?}

>>>>

We experimented with a heuristic
of dynamic construction of the $\sigma$ function, adapted from
\citet{DBLP:conf/pldi/GulwaniZ10}. 
\DM{Cette heuristique, je croyais que c'était nous qui l'avions trouvée,
et qu'elle n'était pas dans Gulwani?}
\JH{Non. Gulwani fait la même chose en un peu plus complique (voir commentaire
suivant)}
We start with one single
disjunct per control point ($m=1$) and an empty map $\sigma$.
$M$ denotes an upper bound on the number of disjuncts per control point.
$\sigma(j,k)$ is then constructed on the fly only when needed.
When $\sigma(j,k)$ needs to be computed, we first compute the image of the
abstract value $X_k$ by the path indexed by $j$, and try to find an existing
disjunct of index $k'$ so that the least upper bound of the two abstract values is exactly their union (this can be tested using SMT-solving).
If such an index exists, then we set $\sigma(j,k) = k'$.
Otherwise:
\begin{itemize}[itemsep=-1\parsep,topsep=0pt]
	\item if $m < M$, we increase $m$ by $1$ and define $\sigma(j,k) = m$
	\item if $m = M$, we define $\sigma(j,k) = M$ 
		\JH{A la place, Gulwani cherche l'invariant qui a le plus de contraintes
		d'egalite en commun pour definir son $sigma$. Nous, on le met dans M, pour simplifier}
\end{itemize}

@@

\DM{Julien: pour la guided analysis, a priori ça arrive souvent cette histoire d'edge pas prise car non encore prenable?}
\JH{Oui, ca arrive regulierement}
In the event that a node is reachable only by a single control-flow edge (which may occur because of dead code, or during the first phases of guided static analysis), the $\phi$ operation reduces to a copy of the values flowing from that edge. In this case, our tool just propagates symbolic values through the predecessor node, without introducing $\phi$-variables.

@@

Statistically, the domain of convex polyhedra gives the better results, but
commonly yields greater invariants than the domains of octagons/intervals;
this is a known weakness of its widening operator.
\DM{Ici on pourrait citer Monniaux+Le Guen si c'etait publie!}
The Octagon domain appears to be much better than intervals; this is
unsurprising since in most programs and libraries, bounds on loop indices are
non constant: they depend on some parameters (array sizes etc.).

@@

The analysis of linear equalities (PKEQ) performs very fast compared
to other abstract domains, but yields very imprecise invariants:
it only detects relations of the form $\sum_i a_i x_i = C$ where $a_i$ and $C$ are constants.
The comparison between the domain of convex polyhedra and linear congruences
shows the percentage of points where the use of congruences refines the
invariant. This is the case for $1.04\%$ of the control points in $P_R$.
\DM{Je croyais qu'on ne pouvait pas utiliser PPLGRID?}
\JH{On peut. C'est PKGRID qui ne fonctionne pas}
